Quiz:

1. Three properties of vector: length, type, name?
2. Type: double, integer, character, factor ; Rare Types: date? and ?
3. Attributes are named informations set to an object? Get them and set thme by 'attr' function?
4. List different from atomic vector because each element in list can be of different type + can store other objects, same for matrix and data frames
5. A list that is a matrix, no. Data frame with column a matrix no.

## Vectors ##

Properties: **Type**, **Length**, **Attributes**

### Atomic vectors ###

Types: logical, integer, double and character

NA is logical vector of length 1, adapted to vector type
Example:
```{r}
a = c(1, 2, 3, 4, NA)
b = NA

typeof(a[5])
#> [1] "double"
typeof(b)
#> [1] "logical"
``` 
**|!|** all elements of vector must be the same type, otherwise **coerced** to be the same type

### Lists ###

**recursive** vectors

complex objects, used to create `lm` and `data.frame` objects

### Exercises ###

1. Six types of atomic vector: **logical**, **character**, **double**, **integer**, **raw** and **complex**
2. is.vector() and is.numeric() does not test if object a vector, returns TRUE only if object is a vector without name
3.
```{r}
c(1, FALSE)
#> [1] 1 0
c("a", 1)
#> [1] "a" "1"
c(list(1), "a")
#> coerces to two lists
c(TRUE, 1L)
#> [1] 1 1
``` 
4. `unlist()` flatten the list while `as.vector()` doesn't
5. `1 == "1"` `TRUE` because of coercion of numbers to character
Same for `-1 < FALSE` coercion of logical to integer
While in `"one" < 2` `"one"` can't be coerced to integer
6. Because logical vectors can coerce to any types

## Attributes ##

Attributes = named list of meta-data

Attributes are **lost** when modifying vector, except **names**, **dimensions** and **class**.
Those attributes have accession function `names(x)`, `dim(x)` and `class(x)`

### Names ###

Naming a vector can be done in 3 ways
```{r}
# At creation
x <- c(a = 0, b = 1, c = 2)

# Modifying it in place
x <- 1:3
names(x) = c("a", "b", "c")

# Creating a modified copy

x <- setNames(1:3, c("a", "b", "c"))
```
new vector without names with `unname(x)` or in place `names(x) <- NULL`

### Factors ###

Integer vector with attributes `class()` and `levels()`

```{r}
x <- factor(c("a", "b", "b", "a"))
x
attributes(x)
```

**|!|** Factors are integers, be careful when using string methods on them (e.g. `gsub()`) they can be coerced to strings, some others will coerce them as integer

### Exercises ###

1. comment is a special attribute assigned using function `comment(x)`, not printed by default
2. Reversing the levels of a factor changes the order in which they are picked, changed the corresponding integer to each level
```{r}
letters

f1 = factor(letters)

levels(f1)

levels(f1) <- rev(levels(f1))

f1

ex = factor(c("a","a", "b", "c"), levels = f1)
ex
ex = factor(ex, levels = rev(f1))
ex
```

3. `f2` has levels in alphabetical order while having the list in reverse orders
`f3` has list in order while levels in reverse order

## Matrices and arrays ##

Arrays and matrices are atomic vectors with attributes `dim()`

matrices: specify dimensions using two integers
arrays: specify dimensions using a vector of integers

`dim()` can also be set on list to make list-matrices or list-arrays

### Exercises ###

1. using `dim()` on vectors return `r NULL`
2. `is.array()` will return `r TRUE`
3. The three objects are similar but with different dimension structure

## Data frames ##

data frame are list of same length vectors

**|!|** do not `cbind` vectors combine them in a data frame!

### Exercises ###

1. A data frame possesses 3 attributes `names`, `row.names` and `class`
2. `as.matrix()` applied to a data frame coerces everything to character
3. it is possible to have a data frame with both 0 rows and 0 columns as:
`r data.frame()` 
