# Notes on Subsetting #

## Quiz ##

1. subsetting a vector with positive integers -> get element at given position
with negative integers -> take element out
logical vector -> return either the entire vector or `integer(0)`
character vector -> `r NA`

2. list
[ -> get first element with attribute returns list
[[ -> returns vector, name or number
$ -> returns vector has to select by name

3. Don't know

4. if x matrix:
`x[] <- 0` assign 0 to all cells in matrix
`x <- 0` will assign 0 to x

5. Don't know

## Data types ##

subset with `[`
positive integer -> element with given index
negative integer -> omit element
logical vector -> select element where `r TRUE`, `r NA` returns a NA in place of element
character vectors -> if vector is named

### Exercises ###

1.
```{r}
mtcars[mtcars$cyl == 4, ]  # Missing = when selecting rows
mtcars[-(1:4), ]  # Missing parenthesis around selecting vector
mtcars[mtcars$cyl <= 5,]  # Missing comma for selecting all columns
mtcars[mtcars$cyl == 4 | mtcars$cyl == 6,]  # OR operator is not enough
```

2. `x <- 1:5; x[NA]` behaves differently from `x[NA_real_]` because `NA` is a logical value:
```{r}
x = 1:5
x[NA]
str(x[NA])  # Use logical value on vector
str(NA)  # Logical value
str(NA_real_)  # Real number
str(x[NA_real_])  # Coerce real number in to integer and indexed column with number does not exist
```

3. `upper.tri()` returns TRUE in the upper triangle of the matrix, to subset cells in matrix no need additional subsetting to describe its behaviours.

4. `mtcars[1:20]` returns an error because there a less than 20 columns in data frame, data frames are considered as list of list when not using the comma.
Thus mtcars[1] returns the first column of mtcars.

5.
```{r}
home.diag = function(x){
	# Test if object is a matrix
	if(!is.matrix(x)){
	stop("Object is not a matrix!")
	}
	nr = nrow(x)
	nc = ncol(x)
	if(nr != nc){
	stop("Can't extract diagonal of a non square matrix")
	}
	vect = c()
	for(i in 1:nc){
	vect = c(vect, m[i,i])
	}
	return(vect)
}

m = matrix(1:16, ncol = 4, nrow = 4)

diag(m)

home.diag(m)
```

6. It replaces all NA values in data frame by 0. `is.na(df)` creates a vector of logical to show where are NA values in data frame. Then we are selecting those cells, and assign them 0 automatically coerced as a vector.

## Subsetting operators ##

### Simplyfing vs. preserving ###
Difference between simplyfing and preserving

simplyfing: get the simplest object no name, only values
preserving: get same type of object, with names

### $ ###

`$` is a shorthand to `[[, exact = F]]`
**|!|** it does partial matching

### Missing indices ###

[[ and [ differs from missing or out of bound indices

### Exercises ###

1.
```{r}

mod = lm(mpg ~ wt, data = mtcars)

mod$df.residual  # Residual degrees of freedom

summary(mod)$r.squared  # get R squared
```

## Subsetting and assignment ##

### Exercises ###

1. Would randomly permute using `sample` function for data frame `df` it would be `df[, sample(ncol(df))]`
2. `df[sample(nrow(df), m), ]` if sample had to be contiguous then sample a single number between 1 and number of rows of df minus length of sample, select this number + m rows
3. For data frame `df` `ordered.col = order(colnames(df))` call `df[, ordered.col]`
