# Notes on Functions #

### Quiz ###
1. Three components of function? arguments, name, and execution?
**Body**, **arguments** and **environment**

2. 11 -> f1(1) returns a function and `f1(1)()` executes the function that adds 10 to x
**TRUE**

3. `2 * 3 + 1`
**TRUE**

4. `mean(x = c(1:10, NA), na.rm = TRUE)`
**TRUE**

5. No it does not because `b` isn't used
**TRUE**

6. Don't know
see **infix**

7. Don't know
use **`on.exit()`**

## Function components ##

All functions has **body**, **arguments** and **environment** accessed by `body()`, `formals()` and `environment()`

### Primitive functions ###
Primitive functions call C code directly using `.Primitive()`

### Exercises ###

1. `is.function()` and `is.primitive()`

2.
```{r}
library(dplyr)

objs = mget(ls("package:base"), inherits = TRUE)
funs = Filter(is.function, objs)

funs.df = sapply(funs, function(x) length(formals(x))) %>%
	as.data.frame(.)
funs.df = funs.df %>%
	mutate(name = row.names(funs.df)) %>%
	rename_(narg = ".")

# Answers question 2.a. it's the scan function that has most argument
filter(funs.df, narg == max(funs.df$narg))

# Answers question 2.b.
filter(funs.df, narg == 0) %>% dim()
```
	c. Can adapt code by `is.primitive` instead of `is.function` in `Filter` function

3. **body**, **args**, **environment**
4. When environment is globalenv

## Lexical scoping ##

### Name masking ###

R look first for variable inside function then a level above, etc.

### Functions vs. variables ###

finding function looks exactly like finding variables, when looking for a function R ignores variables with the same name.

### A fresh start ###

Every time the function is called a new environment is created, each invocation is independent.

### Dynamic lookup ###

R looks for values when the function is **run** not when it was created.

**|!|** It may be a problem if function is run with a spelling error and variable exists in globalEnv -> can know function dependencies using `codetools::findGlobals(f)`

However can make function totally self-contained because R always relies on lexical scoping.

### Exercises ###

1. It returns a vector with first value named "c", because when looking for `c()` ignored the local variable `c`, and the name of the vector can't be a number.

2.
- Name masking
- Fresh start
- Functions vs. variables
- Dynamic Lookup

3. It returns 202

## Every operation is a function call ##

Everything is a function call `[`, `+`, etc.

## Function arguments ##

### Calling functions ###

Arguments can partially match

### Calling a function given a list of arguments ###

Use `do.call()` to use the list of argument to function

### Default and missing arguments ###

Default arguments can be defined using other arguments.

### Lazy evaluation ###

Function arguments are only evaluated when used.

`force()` to ensure an argument is evaluated.

Unevaluated arguments are called **promises**

### `...` ###

Arbitrary list of extra-arguments, makes functions flexibles but error prone.

### Exercises ###

1.
```{r}
x = sample(c(1:10, NA), 20, TRUE)
y = runif(20, 0, 1)
cor(x, y, use = "pairwise.complete.obs", method = "kendall")
```

2. This function returns 3, because the default argument isn't used because already defined in the previous argument. It illustrates lazy evaluation.

3. This function returns 100. Because z is defined in the function environment, it illustrates dynamic lookup inside the function.

## Special calls ##

### Infix functions ###

Prefix function: name comes before the argument
Infix function: name used between the argument (`+`, `:`, etc.)

Can create custom special functions by surrounding name with `%`

### Exercises ###
1.
```{r}
library(stringr)
library(dplyr)

objs <- mget(ls("package:base"), inherits = TRUE)
funs = Filter(is.function, objs)

f.name = function(x) {
	names(x) %>%
	str_detect(regex("^%.*%$"))
}

funs %>%
	f.name() %>%
	names()
```

2. `%ANYCHARACTER%`
3.
```{r}
`%xor%` = function(a, b) {
	if(a!=b){
	TRUE
	} else {
	FALSE
	}
}

a = T
b = F
a %xor% b
```

4.
```{r}
# Infix version of intersect()
`%I%` = function(a, b) {
	vect = c()
	for(i in a) {
		if(i %in% b) {
		vect = c(i, vect)
		}
	}
	sort(vect)
	}

1:10 %I% 4:6

# Infix version of union()
`%U%` = function(x, y) {
	union(x, y)
	}

# Infix version of setdiff()
`%D%` = function(x, y) {
	setdiff(x, y)
	}	
```

5.
```{r}
library(pryr)

`random<-` = function(vect, value) {
	vect[sample(1:length(vect), 1)] = value
	vect
	}

x = 1:10
address(x)

random(x) <- 25
x

address(x)
```

## Return values ##

**Pure fonctions** don't impact workspace

Functions can return invisible values using `invisible()` can wrap functions into () to force them to display

### On exit ###

Can be used to execute code when function exits regardless of the status

### Exercises ###

1.

2. `detach()`. Can save `options()` or `par()` results in list and reload them doing `options(my_list)`

3.
```{r}
graph_device = function(code) {
	dev.new()
	on.exit(dev.off())
	force(code)
	}
```

4. They differ because `capture.output()` is less modular, but handle more exceptions then `capture.output2()`. Also `capture.output()` can save results to a file, while `capture.output2()` can't

